Some Graphical Realizations of Two-Row Specht Modules of Iwahori-Hecke Algebras of the Symmetric Group

### 2019-07-28 tex/Drafts/Draft-1.tex
Literature review

  \documentclass{amsart}   
  \usepackage{../RepStyle} 
  \usepackage{../NatMacros}

@@

  \tableofcontents

@@

\section{Introduction} 
  Let $S_{2n+r}$ be the symmetric group on $2n+r$ indices with $2n + r \geq 2$, let $\SH = \SH_{k,q}(S_{2n+r})$ be the corresponding Hecke algebra over field $k$ with parameter $q \in k^\times$ having square root $q^{1/2}$, and let $\cbr{T_i}$ be the reflections generating $\SH$.
  Let $\brk{m}_q = 1 + q + \dots + q^{m-1}$ be the $q$-number of $m$. 
  Let $e$ be the smallest positive integer such that $\brk{e}_q = 0$, and set $e = \infty$ if no such integer exists.
  Either $q = 1$ and $e$ is the characteristic of $k$ (with $0$ replaced by $\infty$), or $q \neq 1$ and $q$ is a primitive $e$th root of unity.

  {\color{orange} Literature review and motivation goes here.}

  \normalsize
  Throughout the text, we will refer to partitions of $2n + r$;
  identify each partition with a tuple $\lambda = (\lambda_1^{a_1},\dots,\lambda_l^{a_l})$ having $\lambda_i > \lambda_{i+1}$, $a_i > 0$, and $\sum_i a_i\lambda_i = 2n + r$.
  Identify each of these with a subset $\brk \lambda \subset \NN^2$ as defined in Kleshchev, and define $\lambda(i) = (\lambda_1^{a_1},\dots,\lambda_{i-1}^{a_{i-1}},\lambda_i^{a_i - 1},\lambda_i-1,\lambda_{i+1}^{a_{i+1}},\dots,\lambda_l^{a_l})$ to be the partition with the $i$th row removed.

  Fixing some partition $\lambda$, for $1 \leq i \leq j \leq l$, let $\beta(i,j)$ be the hook length
  \[
    \beta(i,j) = \lambda_i - \lambda_j + \sum_{t = i}^j a_t.
  \]
  Then, adopting Kleshchev's terminology, $j$ is normal in $\lambda$ if $\beta(i,j) \not\equiv 0 \pmod e$ for all $i < j$, and $j$ is good if it is the largest normal number (these are stronger conditions than generally necessary).

  Let $S^{(n+r,n)}$ be the Specht module corresponding to the young diagram with two rows with height difference $r$, and let $D^{(n+r,n)}$ be the corresponding irreducible quotient.
  The purpose of this writing is to characterize these representation via an isomorphism with two graphical representations of $\SH$.

@@

  Note that the representations $M_{0 + r}^r$ and $S^{(r)}$ are isomorphic to the sign representation; 
  we will prove that $M_{2n+r}^r$ and $S^{(n+r,n)}$ are isomorphic as representations in the case that $e > n$ and $S^{(n+r,n)}$ is irreducible.
  Note that, when $r = 0$, these have the same dimension given by the $n$th Catalan number $C_n$.

>>>>

  \documentclass{amsart}   
  \usepackage{../RepStyle}

  % Colors links
  \hypersetup{
  colorlinks,
  citecolor=blue,
  linkcolor=black!30!magenta,
  urlcolor=blue}

  \usepackage{../NatMacros}

@@

  \begingroup
  \hypersetup{linkcolor=black} % no colored hyperlinks in toc
  \tableofcontents
  \endgroup

@@

\section{Introduction} 
  Let $S_{2n+r}$ be the symmetric group on $2n+r$ indices with $2n + r \geq 2$, let $\SH := \SH_{k,q}(S_{2n+r})$ be the corresponding Hecke algebra over field $k$ with parameter $q \in k^\times$ having square root $q^{1/2}$, and let $\cbr{T_i}$ be the reflections generating $\SH$.

  Let $\brk{m}_q = 1 + q + \dots + q^{m-1}$ be the $q$-number of $m$. 
  Let $e$ be the smallest positive integer such that $\brk{e}_q = 0$, and set $e = \infty$ if no such integer exists.
  Either $q = 1$ and $e$ is the characteristic of $k$ (with $0$ replaced by $\infty$), or $q \neq 1$ and $q$ is a primitive $e$th root of unity.

  \vspace{5pt}
  When $q = 1$, the Hecke algebra $\SH$ is isomorphic to the group algebra $k\brk{S_{2n+r}}$.
  For $q$ a prime power, let $G := \GL_{2n+r}(q)$ be the general linear group over a field with $q$ elements, let $B := B(q)$ be the Borel subgroup of $G$, and let $1_B$ be the trivial representation of $k\brk{B}$.
  Then, it is a well known result that the endomorphism algebra $\End_{k\brk{G}} \prn{\Ind_B^G(1_B)}$ is isomorphic to the Hecke algebra $\SH$.\cite{Mathas-book}
  Hence the representation theory of the Hecke algebra generalizes the representation theory of the symmetric group, and informs the representation theory of the general linear group.

  It is a classical result that $\SH$ is semisimple precisely when $e > 2n + r$, in which case the irreducible representations of $\SH$ are given by \emph{Specht modules} $S^\lambda$, which are indexed by the partitions $\lambda$ of $2n + r$.
  Further, $\SH$ admits a cellular basis with cell modules given by $S^\lambda$.
  In particular, these admit quotients $D^\lambda$ such that the set $\cbr{D^\lambda \mid D^\lambda \neq 0, \lambda \vdash n}$ is a complete set of pairwise-nonisomorphic irreducible $\SH$-modules.
  This set is indexed by the partitions $\lambda \vdash n$ which are $e$-restricted.\cite{Murphy1,Murphy2}

  These representations $D^\lambda$ are well-defined, but many properties of them are unknown.
  For instance, the dimension of these modules is unknown outside of some special cases.\cite{Mathas-book}
  However, there does exists an algorithm due to Lascoux, Leclerc, and Thibon (verified by Ariki) which produces the decomposition matrices of the Iwahori-Hecke Algebra $\SH_{\CC,q}(S_{2n + r})$ for $q$ an $e$th root of unity.\cite{Ariki,LLT}

  Further, there are theorems, called \emph{modular branching theorems}, characterizing the socle and semisimplicity of the restriction $\Res D^\lambda$ to a subalgebra generated by $2n + r - 2$ simple transpositions.
  These are originally due to Kleshchev in the group algebra case, and were generalized to the Hecke algebra by Brundan.\cite{Brundan,Kleshchev}

  These constructions tend to be well-defined but complicated and often intractable.
  We aim to give simple realizations of some particular $S^\lambda$ and $D^\lambda$ which expose this structure while being intuitive and simple to manipulate.

  \vspace{5pt}
  Throughout the text, we will refer to partitions of $2n + r$;
  identify each partition with a tuple $\lambda = (\lambda_1^{a_1},\dots,\lambda_l^{a_l})$ having $\lambda_i > \lambda_{i+1}$, $a_i > 0$, and $\sum_i a_i\lambda_i = 2n + r$.
  Identify each of these with a subset $\brk \lambda \subset \NN^2$ as is standard, and define $\lambda(i) = (\lambda_1^{a_1},\dots,\lambda_{i-1}^{a_{i-1}},\lambda_i^{a_i - 1},\lambda_i-1,\lambda_{i+1}^{a_{i+1}},\dots,\lambda_l^{a_l})$ to be the partition with the $i$th row removed.\cite{Kleshchev}

  Note that we follow the convention of Murphy and Kleshchev, which is dual to the conventions of Dipper, James, and Mathas;
  one may translate our results to the latter convention by transposing all partitions.\cite{Kleshchev,Mathas-book,Murphy1}

  Fixing some partition $\lambda$, for $1 \leq i \leq j \leq l$, let $\beta_\lambda(i,j)$ and $\gamma_\lambda$ be the quantities
  \begin{align*}
    \beta_\lambda(i,j) &= \lambda_i - \lambda_j + \sum_{t = i}^j a_t\\
    \gamma_\lambda(i,j) &= \lambda_i - \lambda_j + \sum_{t = i+1}^j a_t.
  \end{align*}
  Results due to Kleshchev and Brundan refer to \emph{normal} and \emph{good} numbers;
  note that 1 is always normal, and that $j$ is normal when $\beta_\lambda(i,j) \not\equiv 0 \pmod e$ for all $i \leq j$.
  Recall that $j$ is good iff $j$ is normal and $\gamma_\lambda(j,j') \not\equiv 0 \pmod e$ for all $j' \geq j$ normal.\cite{Brundan,Kleshchev}
  These are the combinatorial properties necessary to state and use branching theorems.

  Throughout this paper, we analyze the \emph{two-row partitions} $(n+r,n) \vdash 2n + r$ and their corresponding modules $S^{(n+r,n)}$ and $D^{(n+r,n)}$.

@@

  Note that the representations $M_{0 + r}^r$ and $S^{(r)}$ are isomorphic to the sign representation; 
  we will prove that $M_{2n+r}^r$ and $S^{(n+r,n)}$ are isomorphic as representations in the case that $e > n + r + 1$.
  Additionally, we will prove that $M_{2n + r}^r$ is irreducible everywhere that $e > n$ and $S^{(n+r,n)}$ is irreducible.
  We will conjecture that $M_{2n + r}^r \simeq S^{(n+r,n)}$ with $e$ unrestricted as well.
